#V2
adding confidence from bootstrap

# Get node trait confidence
```{bash, eval =F, engine.opts = '-l'}
conda activate nextstrain

treetime mugration --tree results/tree.nwk --states data/metadata.csv --attribute region --confidence --outdir mugration/region

```

```{r}
#phylo config
pacman::p_load(treeio, ggtree, tidytree, ape, Biostrings, #phylo
               tidyverse, jsonlite, purrr)

rm(list = ls())
```


#### * Import Tree and mdata

```{r}
#import tree from nextstrain
tree = read.tree("results/tree.nwk")
mdata = read.csv("data/metadata.csv")
fasta = readDNAStringSet("data/sequences.fasta")
node_date0 = jsonlite::read_json("results/branch_lengths.json")

#From Mugration
confidence0 = read.csv("mugration/region/confidence.csv")
mugration_CO0 = read.csv("mugration/region/confidence.csv")
mugration_CO_key0 = read_delim("results/traitsregion.mugration_model.txt", delim = "\t")
```

CHECK IF NAMES MATCH
```{r}

t2 = data.frame(f = sort(names(fasta)), m = sort(mdata$strain)) %>%
  mutate(same = f==m) %>%
  filter(same == "FALSE")

if(nrow(t2) == 0) {
  rm(t2)
}
```



Get Inferred Dates for the Nodes
```{r}

node_index = which(grepl("^NODE", names(node_date0$nodes)))

node_date = node_date0$nodes[node_index]

inferred_dates = lapply(node_date, function(node) node$date)

inferred_dates = as.data.frame(inferred_dates)

inferred_dates = data.frame(t(inferred_dates))

inferred_dates$node = row.names(inferred_dates)

row.names(inferred_dates) = NULL

inferred_dates = inferred_dates %>%
  dplyr::rename("inferred_date" = `t.inferred_dates.`) %>%
  mutate(inferred_date = as.Date(inferred_date))

```



# MDATA_PHYLO:
desc: phylogenetic data (clade, trait confidence, time) & Mdata merge
```{r}
mdata_phylo = mdata

```

# MUGRATION FUNCTION
```{r}

mugration_clean_fun = function(discreet_state, #discrete states from your mdata to help slice the mugration_key
                               mugration_key,  #mugration key output from the mugration function from timetree
                               mugration_data, #mugration matrix data from the mugration function from timetree
                               col_name){      #new colname for your discrete state of inferest
  
  n = length(unique(discreet_state)) #for slice

mugration_key = mugration_key %>%
  dplyr::slice(1:n) %>%
  separate(`Map from character to field name`, into = c("key", "inferred_state"), sep = ":") %>%
  mutate(key = row_number())


colnames(mugration_data) <- as.character(c("node",1:n))

mugration_data %>%
  pivot_longer(cols = -node, names_to = "key", values_to = "confidence") %>%
  mutate(key = as.integer(key)) %>%
  group_by(node) %>%
  filter(confidence == max(confidence)) %>%
  left_join(mugration_key, by = "key") %>%
  mutate(inferred_state = gsub("\t", "", inferred_state)) %>%
  dplyr::select(-key) %>%
    dplyr::rename(!!col_name := inferred_state)
}
```

```{r}

mugration_CO = mugration_clean_fun(discreet_state = mdata$region,
                                     mugration_key = mugration_CO_key0,
                                     mugration_data = mugration_CO0,
                                     col_name = "inferred_state")



mugration_CO = mugration_CO %>%
  dplyr::rename(confidence_state = confidence)

```


#PLOT TREE
# convert tree to df add mdata convert back to tree
```{r}

df_tree = tidytree::as_tibble(tree)

other_trait = c("Brazil", "Canada", "Colombia", "Israel", "Mexico")

df_tree2 = df_tree %>%
           left_join(mugration_CO, by = c("label" = "node")) %>%
           left_join(inferred_dates, by = c("label" = "node")) %>%
           left_join(mdata, by = c("label" = "strain")) %>%
           mutate(tip = !str_detect(label, "^NODE")) %>%
           mutate(inferred_date = dplyr::if_else(is.na(inferred_date),
                                         as.Date(date),
                                         inferred_date)) %>%
           mutate(`Inferred State` = if_else(inferred_state %in% other_trait, 
                                             "Other",
                                             inferred_state))

tree2 = as.treedata(df_tree2)

p_tree = ggtree(tree2) +
  geom_nodepoint(aes(color = `Inferred State`), alpha = 0.7, size = 2.5) +
 # geom_text(aes(label=node), ,size = 2, hjust=-.3) +
  scale_color_brewer(palette = "Dark2") +
  #geom_tippoint(aes(color = inferred_state), size = 1) +
  # scale_color_manual(values = CO_col) +
  geom_treescale()

p_tree

ggsave(plot = p_tree, filename = "data_output/p_tree.png", height = 10, width = 6)
```


#starting from node method 2
```{r}
tree_data = df_tree2
trait = "CO"

# F I L T E R
#get the node numbers for your inferred_state trait of interest
intro_node = tree_data %>% 
  filter(tip == F) %>% #keep only nodes
  filter(confidence_state  >= 0.95) %>% #V2 keep only nodes that have confidence > 0.95
  filter(inferred_state == trait) #keep only nodes that are trait of interest

#extract the node numbers from the dataframe
intro_node_num = intro_node$node

#extract the inferred date to filter to the intro node from the tip 
intro_node_date = intro_node$inferred_date

#get list of nodes of parents a of your node of inferred state of interest
intro_node_list <- map(intro_node_num, ~parent(tree_data, .x))

#add parent nodes inferred state and inferred date to a dataframe for filtering
parent_node = lapply(intro_node_list, function(df) df[, c("node", "inferred_state", "inferred_date")])

#remove nodes where parent interred state is the same as node of interest. AKA keep nodes with a change in the inferred_state aka introduction
parent_node = do.call(rbind, parent_node) %>% 
  dplyr::rename("parent_node" = node,
                "parent_inferred_state" = inferred_state,
                "parent_inferred_date" = inferred_date) %>%
  mutate(intro_node_num = intro_node_num,
         intro_node_date = intro_node_date) %>%
  filter(parent_inferred_state != trait)

#get numbers in a list for the for loop
node_num = parent_node$intro_node_num

#get all the j tip sequences that are in the intro nodes i filter for above
intro_df_list = map(node_num, ~offspring(tree_data, .x, tiponly = T))
                    

#get add the intro node i of the j tips
intro_df_list2 = list()

for(i in 1:(length(intro_df_list))){
  intro_df_list2[[i]]= cbind(intro_df_list[[i]], parent_node[i, c("intro_node_num", "intro_node_date")])
}


#intro_df_list2 = sample(intro_df_list2, 3)

```

FOR LOOP TO FILTER TIPS WITH K ANCESTORS NOT EQUAL TO THE INFERRED STATE
```{r}
tree_data = df_tree2
trait = "CO"

intro_df_list3 = list()

for(i in seq_along(intro_df_list2)){
  
#get node date for the intro node i  
node_date = unique(intro_df_list2[[i]]$intro_node_date)
  
#get the number of tip j node for offspring of node i

#j = 1

tip_ij = as.numeric(intro_df_list2[[i]][,"node"])

  tip_ij_anc=  tip_ij %>%
    map(~ancestor(tree_data, .x))  %>% 
    map(~filter(.x, inferred_date >= node_date))

#add the name of tip_ij to match it with its ancestor nodes 
for(j in 1:length(tip_ij)) {

  tip_ij_anc[[j]] = mutate(tip_ij_anc[[j]], tip_ij = tip_ij[j])
}

#function for true false do all of the inferred state equal the inferred state of interest?
all_ij_trait_ancestor = function(df, x){
  all(grepl(trait, df$inferred_state))
}

#only keep the tips with ancestor nodes that are all the inferred state of interest indicating it is a continued transmission and didn't leave that state
tip_ij_anc2 = keep(tip_ij_anc, all_ij_trait_ancestor)

#extract the tip_ij whos ancestor k were all trait of interest (FoCo)
tip_ij_keep = map_dbl(tip_ij_anc2, ~(unique(.x$tip_ij)))

intro_df_list3[[i]] = intro_df_list2[[i]] %>%
  mutate(tip_keep = if_else(node %in% tip_ij_keep,T,F))

}

intro_GT2 = map_df(intro_df_list3, ~cbind(.x)) 
                    
map_dbl(intro_df_list3, ~filter(.x, tip_keep == T) %>% nrow()) 
```

Get single tips of trait intro 
```{r}
#get the node numbers for your inferred_state trait of interest
intro_node = tree_data %>% 
  filter(tip == T) %>% #changed from above
  filter(inferred_state == trait)

#extract the node numbers from the dataframe
intro_node_num = intro_node$node

#extract the inferred date to filter to the intro node from the tip 
intro_node_date = intro_node$inferred_date

#get list of nodes of parents a of your node of inferred state of interest
intro_node_list <- map(intro_node_num, ~parent(tree_data, .x))

#add parent nodes inferred state and inferred date to a dataframe for filtering
parent_node = lapply(intro_node_list, function(df) df[, c("node", "inferred_state", "inferred_date")])

#remove nodes where parent interred state is the same as node of interest. AKA keep nodes with a change in the inferred_state aka introduction
parent_node = do.call(rbind, parent_node) %>% 
  dplyr::rename("parent_node" = node,
                "parent_inferred_state" = inferred_state,
                "parent_inferred_date" = inferred_date) %>%
  mutate(intro_node_num = intro_node_num,
         intro_node_date = intro_node_date) %>%
  filter(parent_inferred_state != trait)


node_num = parent_node$intro_node_num

single_intro = tree_data %>% 
  filter(node %in% node_num) %>%
  mutate(intro_node_num = node,
         intro_node_date = ymd(date),
         tip_keep = T)
```


# COMBINE INTROS ADD DATA TO FULL TREE
```{r}
intro_samples = rbind(single_intro, intro_GT2) %>%
  filter(tip_keep == T) 

t = intro_samples %>%
  dplyr::select(node, intro_node_num, intro_node_date, tip_keep)

df_tree3 = df_tree2 %>%
  left_join(t, by = "node") %>%
  mutate(intro_tip_state = if_else(!is.na(tip_keep),
                             state,
                             NA)
         )
```

# Write files
```{r}
saveRDS(intro_samples, "data_output/intro_samples_v1.rds")
saveRDS(intro_samples, "../3_persistence_MODELING/data/intro_samples_v1.rds")
saveRDS(df_tree3, "../3_persistence_MODELING/data/df_tree2_v1.rds")
```



